---
title: "Data Horizon System"
subtitle: "M.T.M. Buddy System Motivation"
author: Muhammed Tahir Muhammed
theme: cyborg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(pacman)) install.packages("pacman")
```

# General Social Survey (GSS) data Analysis

Loading the necessary packages using package manager (pacman). Pacman plays an important role in loading packages into the library. It major due is to avoid double installation of a given package.

```{r}
pacman::p_load(tidyverse, plotly, dslabs, esquisse, reactable)
```

The packages above play inportant roles in this analysis. Esquisse via an esquisser enables one to manually create visualization using point and click approach. Plotly enables us to interact with our charts by hovering through.
:::{.panel-tabset}
## The Dataset

### The GSS dataset

Loading the general social survey (gss) dataset into R.

```{r}
reactable(gss_cat)
```
The table above displays the categorical variables in the gss survey.

## Mean Age Summary

### The mean age of the marital categories

Under this, we use the 'verbs' "group_by" and "summarize" from the "dplyr" package embedded in the tidyverse package.

```{r}
mean_age_by_marital_group <- gss_cat |> group_by(marital) |> 
  summarise(mean_age = round(mean(age, na.rm = TRUE),1))
reactable(mean_age_by_marital_group)
```

## Visualization (Chart)

### Visualization of the marital group mean age

```{r}
Mean_age <- mean_age_by_marital_group |> ggplot() + aes(x=marital, y=mean_age)+
  geom_col(aes(fill = marital)) +
  labs(
    x = "Marital Status of the Women",
    y = "Mean Age (year)",
    title = "Bar Chart Showing the mean age by marital Status of Women"
  )
ggplotly(Mean_age)
```
Thanks for reading. You can contact me via [here](https://wa.me/2348167631640?text=Hello!%20I%20am%20a%20Statistics%20and%20Data%20Science%20lecturer%20with%20a%20track%20record%20of%20successfully%20teaching%20and%20mentoring%20students%20on%20data%20analytics%20and%20visualization.%20%0ALet's%20discuss%20the%20way%20forward%20to%20make%20you%20a%20guru%20in%20data%20analytics.){target="_blank"} to teach and mentor you at an affordable price.
:::{.panel-tabset}